Helixer: Cross-species Gene Annotation Of Large Eucaryotic Genomes Using Deep Learning

### 2020-06-02 paper/main.tex
very minor

\begin{figure*}[!tpb]
\label{fig:main_results}
\centerline{\includegraphics[width=\textwidth]{images/main_results}}
\caption{Main overall results and other investigations. The top row shows the animal
 data; the bottom row is concerned with the plants. Orange dots are for AUGUSTUS and 
the blue ones are for Helixer. {\bf a)} Swarm - and boxplots showing the Subgenic F1 
scores for all genomes that were not used during training of Helixer, which is 
equivalent to all validation and test genomes. Each dot represents the prediction 
performace for one genome. The median prediction scores are shown by the red line. 
Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}. {\bf b)} 
Scatterplots showing the prediction performances measured in Subgenic F1 by genome 
length. The regression line is shown with a 95\% confidence interval. {\bf c)} The 
average Genic F1 (in red) and basepair wise accuracy (in purple) with respect to 
the position in the 20000 basepair long input sequences. The dashed line shows the 
same without overlapping. Each value is the average performance of a 200 basepair 
long subsection by one of the models of the final ensemble. 
See Section \ref{ssec:overlapping} of the 
Supplements for the effects of overlapping in individual species.}
\end{figure*}

@@

\begin{figure}[hbt]
\label{fig:average_vs_augustus}
\includegraphics[width=8.6cm]{images/cov_examples/average_vs_augustus}  % should be 8.6 when floats are sorted out right
\caption{Fraction of bp with color-indicated {\bf a)} coverage and {\bf b)} spliced coverage of genomic positions 
broken down by the confusion matrix of AUGUSTUS' and Helixer's predictions. Categories 
are only displayed if they can be meaningfully compared by examining a) coverage or b) spliced
coverage. The displayed fractions are the averages of the individual fractions for the 
six RNAseq-evaluation species. The left-most bars show cases where the two tools agree, 
while the remaining bars show paired conflicts. ``Ntrn" stand
for intron, and ``IG" stands for intergenic.}
\end{figure}

Coverage and Spliced coverage were broken down by the confusion matrix of both
Helixer vs AUGUSTUS as well as Helixer vs the reference for all genomes (Figure \ref{fig:average_vs_augustus}). % H vs A main, H vs R supp
Where both tools agreee on far left in the figure, coverage and spliced coverage closely matched expectations. 
Specifically for CDS:CDS and UTR:UTR most base pairs had some, and many
had moderate or high coverage. The same pattern was seen for intron:intron and spliced
coverage. Finally, intergenic:intergenic showed only a small fraction of base pairs with 
any of either coverage or spliced coverage. In all conflicts the amount of RNAseq support
fell between the cases where tools agreed, indicating that all options were at least
capable of finding weak-spots in the other annotations.

@@

\section{Discussion}
With Helixer, we introduce a novel, deep-learning based framework for the development 
of more capable tools for gene annotation prediction. 
Helixer outperforms AUGUSTUS on base pair wise metrics and on 
consistency with independent RNASeq data while also predicting cross-species for a wide range of  
genomes with one model.

@@

\section*{Acknowledgements}
HHU HPC-team, deNBI, Ebenh{\"o}h lab for computational (particularly GPU) resources.

\section*{Funding}

This work has been supported by the... \vspace*{-12pt}

>>>>

\begin{figure*}[!tpb]
\label{fig:main_results}
\centerline{\includegraphics[width=\textwidth]{images/main_results}}
\caption{Main overall results and other investigations. The top row shows the animal
 data; the bottom row is concerned with the plants. Orange dots are for AUGUSTUS and 
the blue ones are for Helixer. {\bf a)} Swarm - and boxplots showing the Subgenic F1 
scores for all genomes that were not used during training of Helixer, which is 
equivalent to all validation and test genomes. Each dot represents the prediction 
performace for one genome. The median prediction scores are shown by the red line. 
Details of how AUGUSTUS was used are given in Section \ref{sec:augustus}. {\bf b)} 
Scatterplots showing the prediction performances measured in Subgenic F1 by genome 
length. The regression line is shown with a 95\% confidence interval. {\bf c)} The 
average Genic F1 (in red) and basepair wise accuracy (in purple) with respect to 
the position in the 20,000 basepair long input sequences. The dashed line shows the 
same without overlapping. Each value is the average performance of a 200 basepair 
long subsection by one of the models of the final ensemble. 
See Section \ref{ssec:overlapping} of the 
Supplements for the effects of overlapping in individual species.}
\end{figure*}

@@

\begin{figure}[hbt]
\label{fig:average_vs_augustus}
\includegraphics[width=8.6cm]{images/cov_examples/average_vs_augustus} 
\caption{Fraction of bp with color-indicated {\bf a)} coverage and {\bf b)} spliced coverage of genomic positions 
broken down by the confusion matrix of AUGUSTUS' and Helixer's predictions. Categories 
are only displayed if they can be meaningfully compared by examining a) coverage or b) spliced
coverage. The displayed fractions are the averages of the individual fractions for the 
six RNAseq-evaluation species. The left-most bars show cases where the two tools agree, 
while the remaining bars show paired conflicts. ``Ntrn" stand
for intron, and ``IG" stands for intergenic.}
\end{figure}

Coverage and Spliced coverage were broken down by the confusion matrix of both
Helixer vs AUGUSTUS as well as Helixer vs the reference for all genomes (Figure \ref{fig:average_vs_augustus}). 
Where both tools agreee on far left in the figure, coverage and spliced coverage closely matched expectations. 
Specifically for CDS:CDS and UTR:UTR most base pairs had some, and many
had moderate or high coverage. The same pattern was seen for intron:intron and spliced
coverage. Finally, intergenic:intergenic showed only a small fraction of base pairs with 
any of either coverage or spliced coverage. In all conflicts the amount of RNAseq support
fell between the cases where tools agreed, indicating that all options were at least
capable of finding weak-spots in the other annotations.

@@

\section{Discussion}
With Helixer, we introduce a novel, deep-learning based framework for the development 
of more effective tools for gene annotation prediction. 
Helixer outperforms AUGUSTUS on base pair wise metrics and on 
consistency with independent RNASeq data while also predicting cross-species for a wide range of  
genomes with one model.

@@

\section*{Funding}
This work has been supported by the... \vspace*{-12pt}

\section*{Acknowledgements}
We are grateful for the computational suppport and GPU access provided by
the de.NBI cloud, and by both the Centre for Information and Media Technology (ZIM)
and the institute of Quantitative and Theoretical Biology at Heinrich Heine University.
